A performance assessment of evolutionary algorithms in a volunteer computing system

### 2016-03-14 PPSN2016/paper.Rnw
Reviewing Description of the framework.

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
with a client embedded in the browser via JavaScript. We call this
system {\sf NodIO}. All parts of the framework are free and available
with a free license from
\url{https://github.com/JJ/splash-volunteer}.

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for information 
  encoding, in this case chromosomes and related
  information sent by the clients and responses returned by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a cache data structure
  that is reset when the solution is found. This cache has a finite
  size that erases the oldest chromosomes once it has filled to
  capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness chromosome could only be
  inserted once into the cache. That means that if a particular IP
  sends the fitness 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
  % As I remember we kept the run_number so IPs with old runs
  % did not interfere with newer ones. But you are saying: what if they are super slow 
  % and always PUT when the run is already over?, we want to receive 
  % those individuals? 
  % "That means that if a particular IP
  % sends the fitness 99 into the server, it will be accepted only
  % once."  
  % once per run?
% JJ - once per run or experiment. But we didn't keep that run_number
% to avoid interference, just to keep track
  %  but different individuals could have the same fitness, are we limiting the search? 
% We are actually checking for the same individual, not only the fitness
  %  maybe we could have multiple versions of the pool a la MVCC  :)
% Let's leave that for the future...- JJ
  % I think if we are just trying to limit overlapping of the experiments
  % then PUTs should send the run's seq_number and completly remove it.  
% The client does not keep track of the seq number. Maybe we should do
% just that, but sending back and forth the seq number implies
% refactoring. - JJ
\end{enumerate}

@@

NodIO needs a fitness function to work with; in this case the classical Trap 
function \cite{Ackley1987} has been used; this function runs in the
browser, but the server needs to know when the solution has been
found, so a termination condition is also included in the server, as a configuration
option. In all cases we are using 50 Traps, a problem
that is difficult enough to need the intervention of several users for
an extended amount of minutes.

@@

All experiments are announced in the same way: we use social networks,
mainly Twitter, to announce the start of a new experiment. LinkedIn,
Facebook, and also private groups in Telegram and WhatsApp have been
used too for announcements. We also encourage friends and followers to
put an original post for announcing it, instead of just sharing, in
the way allowed by the platform, the original post. During the running
of the experiments, users are engaged, questions answered, and all
kind of explanations given if requested. In general, announcements
have been made in Spanish, but in some cases English was used
too. This makes for a certainly unrealiable experimentation framework,
but it is a realistic one. The steps in the design of the evolutionary
algorithm will be explained next.

>>>>

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
which client is embedded in the browser via JavaScript. We call this
system {\sf NodIO}. All parts of the framework are free and available
with a free license from
\url{https://github.com/JJ/splash-volunteer}.

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for information 
  encoding, in this case chromosomes and related
  information sent by the clients and responses returned by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a cache data structure
  that is reset when the solution is found. This cache has a finite
  size that erases the oldest chromosomes once it has reached its % Also "it has been filled to its"
  capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness chromosome could only be
  inserted once into the cache. That means that if a particular IP
  sends the fitness 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
  % As I remember we kept the run_number so IPs with old runs
  % did not interfere with newer ones. But you are saying: what if they are super slow 
  % and always PUT when the run is already over?, we want to receive 
  % those individuals? 
  % "That means that if a particular IP
  % sends the fitness 99 into the server, it will be accepted only
  % once."  
  % once per run?
% JJ - once per run or experiment. But we didn't keep that run_number
% to avoid interference, just to keep track
  %  but different individuals could have the same fitness, are we limiting the search? 
% We are actually checking for the same individual, not only the fitness
  %  maybe we could have multiple versions of the pool a la MVCC  :)
% Let's leave that for the future...- JJ
  % I think if we are just trying to limit overlapping of the experiments
  % then PUTs should send the run's seq_number and completly remove it.  
% The client does not keep track of the seq number. Maybe we should do
% just that, but sending back and forth the seq number implies
% refactoring. - JJ
\end{enumerate}

@@

NodIO needs a fitness function to work with; in this case the classical Trap 
function \cite{Ackley1987} has been used; % I don't think this use of the ; is correct
this function runs in the
browser, but the server needs to know when the solution has been
found, so a termination condition is also included in the server, as a configuration
option. In all cases we are using 50 Traps, a problem
that is difficult enough to need the intervention of several users for
an extended amount of minutes.

@@

All experiments are announced in the same way: we use social networks,
mainly Twitter, to announce the start of a new experiment. LinkedIn,
Facebook, and also private groups in Telegram and WhatsApp have been
used for announcements too. We also encourage friends and followers to
put an original post for announcing it, instead of just sharing, in
the way allowed by the platform, the original post. During the running
of the experiments, users are engaged, questions answered, and all
kind of explanations given if requested. In general, announcements
have been made in Spanish, but in some cases English was used
too. This makes for a certainly unrealiable experimentation framework,
but it is a realistic one. The steps in the design of the evolutionary
algorithm will be explained next.
